---
title: "Lab2"
author: "Greg Ceccarelli"
date: "June 30, 2015"
output: pdf_document
---

## Part 1. Multiple Choice (5 points each)

1. 
2. 
3. 
4. 
5. 
6. 
7. 
8. 
9. 
10.

## Part 2. Test Selection (10 points)

Every other year, the General Social Survey collects responses to thousands of questions,
covering a wide variety of topics. You will be using a subset of data from 1993,
including a small number of variables. 

This may be found in the file, GSS.Rdata.

Like any survey, GSS data creates additional concerns that would normally go into a
statistical analysis. Surveys are usually weighted in order to compensate for over- or
under-representation of subgroups. For this lab, however, you will be using unweighted
data, which limits how well your findings generalize to the U.S. population.
For the following problems, select the most appropriate statistical test to answer the
question from the given choices. 


11.

12. 

# Part 3: Data Analysis and Short Answer (50 points )

13. Data Import and Error Checking
a. Examine the “agewed” variable (age when married). What are the value(s)
of agewed, if any, that do not meaningfully correspond to ages?
**RESPONSE:** There are 286 instances where agewed = 0 and 12 instances where agewed = 99. Both of these values seem highly unlikely and do not meaninfully correspond to ages (although it is technically feasible to wed at age 99)

b. Recode any value(s) that do not correspond to age as NA. What is the
mean of the agewed variable?

```{r}
#set up working envrionment
setwd('/Users/ceccarelli/MIDS/DATASCI_W203/Assignments/Labs/Lab 2')
rm( list = ls() )

#load supplied R data file
load("GSS.Rdata")

#view tabulated counts by value
data.frame(table(GSS$agewed))

#replace outlier values
GSS$agewed[GSS$agewed == 0 | GSS$agewed == 99] <- NA 

vars<-c("agewed","age")

data.frame(table(GSS[vars]))


```


14. Checking assumptions
a. Produce a QQ plot for the agewed variable. Using this plot information, is
agewed normal and how do you know?

b. Perform a Shapiro-Wilk test on the agewed variable.
i. What is the null and alternative hypothesis for your test?
ii. What is your p-value, and what is your specific conclusion?

c. What is the variance of agewed for men? What is the variance of agewed
for women?

d. Perform a Levene’s test for the agewed variable grouped by men and
women.
i. What is the null and alternative hypothesis for this test?
ii. What is your p-value, and what is your specific conclusion?

```{r}


```

15.More hypothesis testing

a. Suppose we have a hypothesis that the age of marriage (agewed) in the
population has a mean of exactly 23, with a standard deviation of 5 years

(you should assume this value is correct rather than estimating the standard
deviation from the data). Perform a z-test to analyze this hypothesis.

i. What is the null and alternative hypothesis for this test?
ii. What is your p-value, and what is your specific conclusion?
