---
title: "R Exercise #2"
author: "Greg Ceccarelli"
date: "June 23, 2015"
output: pdf_document
---

*Question 1*

Write a function in R that does the following:
* Takes sample data (e.g., as a vector), mean of the null hypothesis, and population standard deviation, and a Boolean variable indicating whether we want a one-tailed or two-tailed test. Returns p-values for the required test (Null is “population mean is the one that the function is fed”) , as well as a Boolean value showing whether the test passes a conventional 5% level or not.

* Choose a sample size and then use rnorm(n, mean, sd) to generate a random sample and test your function.


```{r}

##import plyr to use rdply
library(plyr)

testfun <- function(data=c(), mu, pop.sd, two.tailed=TRUE, significance=.05) {
      ##computed sample mean
      xbar <- mean(data)
      ##computed sample size
      n <- length(data) 
      ##compute z score
      z <- (xbar - mu) / (pop.sd /sqrt(n))
     
      #modify and compute pvalue depending on type of test required
       if (two.tailed==TRUE) {
      p <- 2 * pnorm(z, lower.tail = FALSE)
      } else if (two.tailed==FALSE & xbar > mu) {
      p <- pnorm((z), lower.tail = FALSE)
      } else if (two.tailed==FALSE & xbar < mu) {
      p <- pnorm((z))
      }
      
      #create boolean value depending on 
      if (p>significance) {
      pass <- FALSE
      } else {
      pass <- TRUE
      }
    
      
    ##return a dataframe to make storing replicated results easy
    df <- data.frame(p, pass)
    colnames(df) <- c("pvalue","pass")
    return(df)    
}

##test function with rnorm
testfun(rnorm(200,197.3,25.6),191,25.6,FALSE,.05)
```

* Use replicate or sapply to generate a sample and do the test multiple, say 1000, times. Plot the histogram of p-values that you are getting when the Null is true. Is your function calculating Type-I errors correctly?

```{r}
##import plyr to use rdply
library(plyr)

##repeat previous example with test function with rnorm using rdply to replicate 100 times
result <- rdply(100, testfun(rnorm(200,197.3,25.6),191,25.6,FALSE,.05))

##save pvalues only when NULL is true (ie, pass==FALSE)
null.true <- result[result$pass==FALSE,]

##plot p-values to understand type 1 error
hist(null.true$pvalue)
```


* Now assume your Null is false. Note: For type-II error calculation, you need a specific assumption about the mean of the population from which the sample is taken. Assume that this mean is one standard deviation above the Null mean.

* Calculate type-II errors both theoretical and by simulation as in step 3. Then increase the sample size and redo this.
What happens to statistical power when you increase sample size?


```{r}
##repeat previous example with test function with rnorm using rdply to replicate 1000 times
result <- rdply(1000, testfun(rnorm(70,2,.25),2.09,.25,FALSE,.05))
result$power <- 1 - (result$pvalue)
##save pvalues only when NULL is FALSE (ie, pass==TRUE)
null.false <- result[result$pass==TRUE,]
##plot p-values to understand type 2 error
hist(null.false$pvalue)
summary(result$pass)


## doubling sample sizes
result <- rdply(1000, testfun(rnorm(140,2,.25),2.09,.25,FALSE,.05))
result$power <- 1 - (result$pvalue)
##save pvalues only when NULL is FALSE (ie, pass==TRUE)
null.false2 <- result[result$pass==TRUE,]
##plot p-values to understand type 2 error
hist(null.false2$pvalue)
summary(result$pass)

```