---
title: "Final"
author: "Greg Ceccarelli"
date: "August 16, 2015"
output: pdf_document
---

## Part 1. Multiple Choice (32 points)

1. 
2. 
3. 
4. 
5. 
6. 
7. 
8. 


## Part 2. Test Selection (24 points)

9. a
10. d
11. b could be c
12. b
13. a
14. b

```{r}

##TO DELETE THIS SECTION, USED TO EXPLORE VARIABLES TO DECIDE ON TESTS

summary(dating$lgbt)
summary(dating$adults_in_household)

anova(dating$lgbt,dating$adults_in_household)

summary(dating$children0_5) 
summary(dating$children6_11)
summary(dating$children12_17)


```

## Part 3: Data Analysis  (44 points )

15. OLS Regression

```{r}
#set up working envrionment
##running on macbook air, set relevant directory
setwd('/Users/ceccarelli/MIDS/DATASCI_W203/Assignments/Labs/Final Exam/')
rm( list = ls() )

#Load relevant packages
library(ggplot2)
library(car)
library(psych)
library(gmodels)
library(MASS)
library(plyr)

#load supplied R data file
dating <- read.csv("Dating.csv",header = TRUE)

#inspect life_quality variable
summary(dating$life_quality)
levels(dating$life_quality)

#wrapper function to easily recode factors
changelevels <- function(f, ...) {
    f <- as.factor(f)
    levels(f) <- list(...)
    f
}

dating$life_quality_cleaned <- changelevels(dating$life_quality, "5"=c("1"), "4"=c("2"), "3"=c("3"),"2"=c("4"),"1"=c("5"),"NA"=c("Don't know","Refused"))

levels(dating$life_quality_cleaned)

##appears NA is a string in dating$life_quality_cleaned, recode to true NA
is.na(dating) <- dating=="NA"

# Can fix the remaining "NA" after fix by recreating the factor 
dating$life_quality_cleaned <- factor(dating$life_quality_cleaned)

#check summary output
summary(dating$life_quality_cleaned)

#convert to numeric to use in regression using ali's method
dating$life_quality_cleaned <- as.numeric(as.character(dating$life_quality_cleaned))

#compute the mean for life quality to answer 15. a
mean(dating$life_quality_cleaned, na.rm =TRUE)

summary(dating$life_quality_cleaned)

hist(dating$life_quality_cleaned)


##review values for dating$years_in_relationship & age
counts_yrs = as.data.frame(table(dating$years_in_relationship))
counts_yrs[with(counts_yrs,order(-Freq)),]

counts_age = as.data.frame(table(dating$age))
counts_age[with(counts_age,order(-Freq)),]

dating$years_in_relationship_cleaned <- dating$years_in_relationship


#in this instance, use plyr instead to mapvalues
dating$years_in_relationship_cleaned <- mapvalues(dating$years_in_relationship_cleaned, from = c(" ","Refused"), to= c("NA", "NA"))

##appears NA is a string in dating$years_in_relationship_cleaned, recode to true NA
is.na(dating) <- dating=="NA"

# Can fix the remaining "NA" after fix by recreating the factor 
dating$years_in_relationship_cleaned <- factor(dating$years_in_relationship_cleaned)

#recheck counts to make sure nothing amiss
counts = as.data.frame(table(dating$years_in_relationship_cleaned))
counts[with(counts,order(-Freq)),]

#check total counts, 59 NA's not included
sum(as.numeric(counts$Freq))
hist(as.numeric(dating$years_in_relationship_cleaned))

#update to be numeric following Alis Guidance
dating$years_in_relationship_cleaned <- as.numeric(as.character(dating$years_in_relationship_cleaned))

##finally, check for spurious values when years_in_relationship exceeds age
dating$years_in_relationship_cleaned.spurious <- as.numeric(as.character(dating$years_in_relationship_cleaned)) > dating$age

table(dating$years_in_relationship_cleaned.spurious)

##One value where this occurs, recode to NA
#View(data.frame(dating[dating[,"years_in_relationship_cleaned.spurious"],]))

is.na(dating$years_in_relationship_cleaned) <- dating$years_in_relationship_cleaned.spurious==TRUE

dating$years_in_relationship_cleaned <- factor(dating$years_in_relationship_cleaned)

#compute the mean for life quality to answer 15. b
mean(dating$years_in_relationship_cleaned, na.rm =TRUE)

##check use_internet 
summary(dating$use_internet)

#remap values
dating$use_internet_cleaned <- mapvalues(dating$use_internet, from = c(" ","Don't know","Refused"), to= c("NA","NA","NA"))

## NA is a string in dating$use_internet_cleaned, recode to true NA
is.na(dating) <- dating=="NA"

# Can fix the remaining "NA" after fix by recreating the factor 
dating$use_internet_cleaned<- factor(dating$use_internet_cleaned)

##check use_internet again
summary(dating$use_internet_cleaned)

#compute lim_rows logical vector for the 3 variables in question
lim_rows = complete.cases(dating$life_quality_cleaned, dating$use_internet_cleaned, dating$years_in_relationship_cleaned)

## just the complete cases, count the rows with nrow to answer 15. c
dating_lim = dating[lim_rows,]
nrow(dating_lim)


lmod1 <- lm(life_quality_cleaned ~ years_in_relationship_cleaned, dating_lim)
plot(dating$life_quality_cleaned ~ dating$years_in_relationship_cleaned)
abline(lmod1)

##plot model to review
plot(lmod1)

summary(lmod1)
#first models coefficient
coef(lmod1)

#fit multivariate ols linear model: life_quality outcome, use_internet
#and years_in relationship as predictors
lmod2 <- lm(life_quality_cleaned ~ years_in_relationship_cleaned + use_internet_cleaned, dating_lim)

summary(lmod2)
coef(lmod2)

# compare the model improvement with anova
anova(lmod1, lmod2)

# also check the AIC
AIC(lmod1)
AIC(lmod2)



```


**RESPONSE:** 
A. What is the mean quality of life in the sample?
3.392921

B. What is the mean of "years in relationship" in the sample?
13.43887

c. How many cases does this leave you with?
1089

D. Fit an OLS Model. What is the slope coefficient you get? Is it
statistically significant? What about practically significant?
The slope coefficient is  0.005217756, meaning that every unit increase
in "years in relationship" increases "life quality" by that amount. It is
statistically significant but given the extremely small R squared, not
practically significant given the very low percent of variance explained.


E. Fit a second OLS model. What is the slope coefficient for use_interne
? Is it statistically significant? What about practically significant?
The slope coefficient for use internet is  0.404609136, meaning that internet users on average had a rating for "life quality" that much higher. In this case it is higher statistically significant with a p value of 5.19e-06. 


F. Compute the F-ratio and associated p-value between your two regressio
models. Assess the improvement from your first model to your second.
Using ANOVA, The F-ratio is 20.975 and is highly statisitically significant. Reduced AIC between model 1 and 2 indicate that we've improved predictive ability and that we have a higher quality model (with improved R squared)


16. Logistic Regression


```{r}
summary(dating$flirted_online)

#remap values
dating$flirted_online_cleaned <- mapvalues(dating$flirted_online, from = c(" ","Don't know","Refused"), to= c("NA","NA","NA"))

## NA is a string in dating$flirted_online_cleaned, recode to true NA
is.na(dating) <- dating=="NA"

# Can fix the remaining "NA" after fix by recreating the factor 
dating$flirted_online_cleaned<- factor(dating$flirted_online_cleaned)

summary(dating$flirted_online_cleaned)


#create dummy variables
contrasts(dating$flirted_online_cleaned)<-contr.treatment(2,base = 1)

levels(dating$flirted_online_cleaned)

summary(dating$flirted_online_cleaned)

dating$flirted_online_cleaned_recode <- ifelse(dating$flirted_online_cleaned=="Yes", 1, 0)

#t=table(dating$flirted_online_cleaned)
#D = data.frame( matrix(t,ncol=2))
#D$num.obs = D[,1]+D[,2]
#D$srate = D[,2]/D$num.obs

##check updated values
summary(dating$flirted_online_cleaned_recode)

##compute mean
mean(dating$flirted_online_cleaned_recode, na.rm=TRUE)

#compute odds for question 16. a
mean(dating$flirted_online_cleaned_recode, na.rm=TRUE)/(1-mean(dating$flirted_online_cleaned_recode, na.rm=TRUE))


#review usr variable
summary(dating$usr)

dating$usr_cleaned <- mapvalues(dating$usr, from = c(" "), to= c("NA"))

## NA is a string in dating$flirted_online_cleaned, recode to true NA
is.na(dating) <- dating=="NA"

# Can fix the remaining "NA" after fix by recreating the factor 
dating$usr_cleaned<- factor(dating$usr_cleaned)

summary(dating$usr_cleaned)

# Only incorporate complete cases
#compute lim_rows logical vector for the 3 variables in question
lim_rows_glm = complete.cases(dating$flirted_online_cleaned_recode, dating$usr_cleaned)

## just the complete cases
dating_lim_glm = dating[lim_rows_glm,]
summary(dating_lim_glm$usr_cleaned)


#create dummy variables, make surburban base case
contrasts(dating$usr_cleaned)<-contr.treatment(3,base = 2)

## Begin with a bivariate logistic regression given 
model1 = glm(flirted_online_cleaned_recode ~ usr_cleaned, data=dating_lim_glm, family=binomial())

#grab AIC criterion
summary(model1)

exp(coef(model1))



#create dummy variables, make surburban base case
contrasts(dating$usr_cleaned)<-contr.treatment(3,base = 2)

dating$usr_cleaned

exp(coef(model1))

```


**RESPONSE:** 
A. What are the odds that a respondent in the sample has flirted online at some point (flirted_online)?
The percentage of folks who flirted online is 0.2072072, thus, the odds that someone would have flirted online is 0.2072072/(1-0.2072072) = .26136. On the otherhand, the odds that someone did not flirt online is (1-0.2072072)/0.2072072 = 3.82

B. Conduct a logistic regression to predict flirted_online as a function
of where a respondent lives (usr). What Akaike Information Criterion (AI
) does your modelhave?
1909.4

C.  how much bigger are the odds that an urban respondent thas flirted
online than the odds that a rural respondent has flirted online? Is this
effect practically significant?
