---
title: "Lab 3"
author: "Greg Ceccarelli"
date: "July 29, 2015"
output: pdf_document
---

## Part 1. Multiple Choice (4 points each)

1. 
2. 
3. 
4. 
5. 
6. 
7. 
8. 



## Part 2. Test Selection (20 points)

9. 

10.

11.

12.

13.


## Part 3: Data Analysis and Short Answer (48 points )

14. Task 1: Conduct a chi-square test to determine if there is an association
between marital status (marital) and political orientation (politics).

```{r}
#set up working envrionment
##running on macbook air, set relevant directory
setwd('/Users/ceccarelli/MIDS/DATASCI_W203/Assignments/Labs/Lab 3/')
rm( list = ls() )

#Load relevant packages
library(ggplot2)
library(car)
library(psych)
library(gmodels)
library(MASS)

#load supplied R data file
load("GSS.Rdata")

#define function to omit row values where only certain columns are NA
data.complete <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

#question 14
#inspect variable to check on categories
summary(GSS$marital)

##appears NA is a string in Marital, recode to true NA
is.na(GSS) <- GSS=="NA"

# Can fix the remaining "NA" after fix by recreating the factor 
GSS$marital <- factor(GSS$marital)

# true NA's now
summary(GSS$marital)

# check politics as well
summary(GSS$politics)

#created NA subsetted dataset to run chi square test on
GSS.mar_pol <- data.complete(GSS[,c("marital","politics")], c("marital","politics")) 


#review tabulated results after removing null
table(GSS.mar_pol)

##leverage CrossTable function to compute Pearson's chisq test 
CrossTable(GSS.mar_pol$marital, GSS.mar_pol$politics, fisher = TRUE, 
           prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE,
           chisq = TRUE, expected = TRUE, sresid = TRUE, format = "SPSS")

# also leverage a different function to compare results
cs <- chisq.test(GSS$marital, GSS$politics)

#since this is a test of indepedence, leverage cramer's v for effect size
cramers_v = function(cs)
{
  cv = sqrt(cs$statistic / (sum(cs$observed) * (min(dim(cs$observed))-1)))
  print.noquote("Cramer's V:")
	return(as.numeric(cv))
}

cramers_v(cs)
```


**RESPONSE:** 
A. What are the null and alternative hypothesis for your test?
We use the chi-sq test to test the relationship between marital status 
The null hypothesis is that marital status and political orientation are
indepedent the alternative hypothesis is that marital status and political
orientation are not indepedent (they are associated)

B. What test statistic and p-value do you get?
The value of the chi-square statistic is 44.2255. 
This value is highly significant (p < .0001), indicating that marital status has
a significant effect on political orientation. 

C. Conduct an effect size calculation for your relationship.
Using cramer's v returns an effect size of 0.08756363.

D. Evaluate your hypothesis in light of your tests of statistical and 
practical significance. What, if anything, can you conclude from your results?
While the results from the chi-square test are statistically significant, 
it is hard to make a good conclusion given the lack of practical significance
(anything under .2 is considered weak). 
Thus while the variable's are associated, their association is not very strong.



15. Task 2: Conduct a Pearson correlation analysis to examine the association
between age when married (agewed) and hours of tv
watched (tvhours).

```{r}
#question 15
#check how linear the variables are
#inspect variable 
describe(GSS$agewed)

#agewed is is not normally distributed and exhibits leptokurtic positive skew
hist(GSS$agewed)

#inspect tvhours
describe(GSS$tvhours)

#tvhours are even more non-normally distributed and have some outliers
hist(GSS$tvhours)

#subset dataset because tvhours variable  cannot exceed 24
GSS.age_tv <- subset(GSS, tvhours <= 24, select = c(agewed,tvhours))

#subset again because 0, 98 and 99 are invalid for agewed
GSS.age_tv <- subset(GSS.age_tv, agewed > 0 & agewed<=90)

#review histogram again
hist(GSS.age_tv$tvhours)
hist(GSS.age_tv$agewed)

# Use a scatterplot to see how linear the relationship looks on the new
# variables
scatterplot(GSS.age_tv$agewed, GSS.age_tv$tvhours)

#check the correlation
cor.results <- cor.test(GSS.age_tv$agewed, GSS.age_tv$tvhours, 
                        use = "complete.obs", method = "pearson")

cor.results

# compute r^2
cor.results$estimate**2

## compute r^2 another way with a correlation matrix
cor(GSS.age_tv[,c("agewed", "tvhours")], use = "complete.obs")**2

```


**RESPONSE:** 
A. What are the null and alternative hypotheses for your test?
The null hypothesis is that the linear relationship between agewed and tvhours
is 0
The alternative hypothesis is that the linear relationship is not 0

B. What test statistic and p-value do you get?
Test statistic of -1.0349 and a p-value of .3009. 

Also, an R value of -.0299. 95 percent confidence interval -0.08654554 and 
0.02681630. R^2 value of 0.0008976593. 

C. Evaluate your hypothesis in light of your tests of statistical and practical
significance. What, if anything, can you conclude from your results?
Given the very high pvalue, there is not enough evidence to reject the null
hypothesis that the linear relationship is 0 (aka there is not one). This is
confirmed via practical signifance and the confidence interval that spans 0.


16. Task 3: Create a new binary/dummy variable, “married”, that denotes whether
an individual is currently married or not currently married. Next, we want to
consider just the subpopulation of 23-year olds in this sample. 
Conduct a Wilcox rank-sum test to determine whether your new “married” variable 
is associated with the number of children (childs) for respondents who are 23
years old.



```{r}

#question 16
#create dummy variable for married using if else
GSS$marital_dummy <- ifelse(GSS$marital == "married",1,0)

#check summary stats for this variable
summary(GSS$marital_dummy)

#create subset for variable
GSS.marital_dummy <- subset(GSS, age == 23, select = c(age,marital_dummy,childs
                                                       ))
#View(GSS.marital_dummy)

##look at mean/proportion of married in new subset
mean(GSS.marital_dummy$marital_dummy)
summary(GSS.marital_dummy)

# We may ask whether the married 23 years olds have more or less children
by(GSS.marital_dummy$childs,GSS.marital_dummy$marital_dummy, mean, na.rm = TRUE)

# Notice that number of children is not at all normal.
qqnorm(GSS.marital_dummy$childs)

# Use the Wilcoxon rank-sum test to compare means
wilcox.test(GSS.marital_dummy$childs ~ GSS.marital_dummy$marital_dummy)

# leverage cohensd function from previous work to compute effect size
cohens_d <- function(x, y) {
  # this function takes two vectors as inputs, and compares
	# their means
	
	# first, compute the pooled standard error
  lx = length(subset(x,!is.na(x)))
  ly = length(subset(y,!is.na(y)))
	# numerator of the pooled variance:
	num = (lx-1)*var(x, na.rm=T) + (ly-1)*var(y, na.rm=T)
	pooled_var = num / (lx + ly - 2) # variance
	pooled_sd = sqrt(pooled_var)
	
	# finally, compute cohen's d
	cd = abs(mean(x, na.rm=T) - mean(y, na.rm=T)) / pooled_sd
	return(cd)
}


#need vectors to plug into cohens d
children_m = GSS.marital_dummy$childs[GSS.marital_dummy$marital_dummy==1]
children_nm = GSS.marital_dummy$childs[GSS.marital_dummy$marital_dummy==0]

# plug them into our cohen's d function
cohens_d(children_m, children_nm)

```



**RESPONSE:** 
A. What is the mean of your new “married” variable among 23-year-olds (e.g., the proportion of cases in the category coded “1”)?
.2857 or 28%

B. What is the null and alternative hypotheses for your test?
The null hypothesis is the mean number of children had by individuals (married or not) is identical
The alternative hypothesis is that the mean number of children had by individuals (married or not) is not identical

C. What test statistic and p-value do you get?
W = 19 and p = .0002656

D. Conduct an effect size calculation for your relationship.
Using cohens d, the effect size is extraordinarily big! 1.976885

E. Evaluate your hypothesis in light of your tests of statistical and practical significance. What, if anything, can you
conclude from your results?
In light of evaluating the extremely significant p value and the large effect size, we can safely conclude that the means
between these two groups are different.




17. Task 4: Conduct an analysis of variance to determine if there is an association between religious affiliation (relig) and
age when married (agewed).



```{r}

#question 17

library(ggplot2)
##inspect religion variable
summary(GSS$relig)

#check out levels
levels(GSS$relig)

#wrapper function to easily recode factors
changelevels <- function(f, ...) {
    f <- as.factor(f)
    levels(f) <- list(...)
    f
}

GSS$relig_clean <- changelevels(GSS$relig, Protestant=c("Protestant"), Catholi=c("Catholic"), Jewish=c("None"),None=c("Non"),Other=c("Other","DK"))

#recheck out levels
levels(GSS$relig_clean)

#remove spurious agewed variables
GSS.age_relig <- subset(GSS, agewed > 0 & agewed<=90)

#remove NA's
GSS.age_relig <- data.complete(GSS.age_relig ,c("agewed","relig_clean"))

# check the agewed variable for normality
hist(GSS.age_relig$agewed)

qqplot = qplot(sample = GSS.age_relig$agewed, stat="qq")
qqplot

# Normality is not great but ANOVA is a robust-test
# also data is interval

# Let's look at the means, by each category and overall
by(GSS.age_relig$agewed, GSS.age_relig$relig_clean, mean, na.rm=T)
mean(GSS.age_relig$agewed, na.rm=T)

# Perform the analysis of variance and check the significance
aovm = aov(agewed ~ relig_clean, GSS.age_relig)
summary(aovm)

# perform a post hoc test to check for pairwise differences in means... use
bonferroni correction
pw = pairwise.t.test(GSS.age_relig$agewed, GSS.age_relig$relig_clean, p.adjus
                     .method = "bonferroni")

pw
#make output clearer using a function
sig_stars = function(p)
{
  stars = symnum(p, na = F, cutpoints = c(0, .001, .01, .05, .1, 1), symbols=c(
    ***","**", "*", ".", " "))
	return( paste(round(p, 3), stars) )
}

# apply our new function to every element in our matrix
t_table = noquote( apply(pw$p.value, c(1,2), sig_stars) ) 
t_table


```


**RESPONSE:** 

A. What is the null and alternative hypotheses for your test?
The null hypothesis is that there are no differences in the mean wed age 
between religious groups
The alternative hypothesis is that there are differences in the mean wed 
age between the religious groups

B. What test statistic and p-value do you get?
F=8.199 and p value 1.6e-06. The overall (omnibus) model is very significant.

C. Are there statistically significant differences between individual pairs of
groups, and if so, how do you know?
Yes, using a post hoc test (pairwise ttest with the bonferroni correction),
shows that there are statistically significant differences in wed age between
Protestant and Catholics (largest significant difference),  Protestant and
Jewish and Protestant and Other groups.


D. Evaluate your hypothesis in light of your tests of statistical and practical
significance. What, if anything, can you conclude from your results?

While there is an overall difference in mean age when wed, the most interesting
statistically significant differences occur between Protestants and all groups
except (None)





